Chr7 hotspot analyses

### 2019-07-08 analysis/paper/Chr5-prelim.Rmd
updated Rmd files for each hotspot

```{r, load_attie}
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
ls()
library(tidyverse)
```

@@

```{r}
Chr5hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
Chr5hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
```

@@

```{r fit1_med}
nlocal <- ncol(Chr5hot_local)
nnonlocal <- ncol(Chr5hot_nonlocal)
out <- list()
out2 <- list()
for (j in 1:nlocal){
  foo <- list()
  foo2 <- list()
  for (i in 1:nnonlocal){
    foo[[i]]<- qtl2mediate::fit1_med(driver = genoprobs$`5`[rownames(genoprobs$`5`) %in% rownames(Chr5hot_local) , , indices[i]], 
                                      # genoprobs at peak position for nonlocal trait
                                      target = Chr5hot_nonlocal[ , i], 
                                      mediator = Chr5hot_local[ , j], 
                                      addcovar = addcovar
         )
    foo2[[i]] <- qtl2mediate:::fit1_med_qtl2(driver = genoprobs$`5`[, , indices[i]],
                                             target = Chr5hot_nonlocal[ , i],
                                             mediator = Chr5hot_local[ , j],
                                             addcovar = addcovar
                                             )
  }
  out[[j]] <- foo
  out2[[j]] <- foo2
}
```

@@

## Reading pleiotropy test results

```{r read_pleio_results}
results_dir <- "~/Box Sync/attie/keller-hotspots/results/pvl-run-9901"
dir(results_dir) -> fns
mylist <- list()
for (i in 1:length(fns)){
  mylist[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
  as_tibble() %>%
  mutate(filename = fns[i]) %>%
  mutate(local_gene_id = stringr::str_split(fns[i], "_")[[1]][[3]], nonlocal_gene_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
  filter(lod == max(lod), trace == "profile1") %>%
  select(local_gene_id, nonlocal_gene_id, lod, filename)
}
mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  group_by(nonlocal_symbol) %>%
  filter(lod_diff == max(lod_diff)) %>%
  arrange(local_symbol, desc(lod_diff))
```

@@

## Are the lods from fit1_med correct?

Yes.

I see that a large number of nonlocal transcripts have `lod_without_med` values below 7.18. While the 7.18 threshold was determined with the mixed model, I still find the large number of transcripts unusual.

First, a given nonlocal transcript should have a single value for `lod_without_med`, regardless of which local transcript it's paired with. Let's verify this.

```{r, check1}
mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>% 
  arrange(nonlocal_symbol)
```

```{r, check_table}
mylist %>%
  bind_rows() %>%
  select(lod) %>%
  mutate(indic = lod > 5) %>%
  select(indic) %>%
  table()
```

@@

>>>>

```{r, load_attie}
hot_chr <- 5
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
ls()
library(tidyverse)
```

@@

```{r}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
```

@@

```{r fit1_med}
nlocal <- ncol(hot_local)
nnonlocal <- ncol(hot_nonlocal)
medfile <- "_cache/Chr5_mediations.rds"
if (!file.exists(medfile)){
  out <- list()
  for (j in 1:nlocal){
  foo <- list()
  for (i in 1:nnonlocal){
    foo[[i]] <- qtl2mediate:::fit1_med_qtl2(driver = genoprobs[[as.character(hot_chr)]][, , indices[i]],
                                             target = hot_nonlocal[ , i],
                                             mediator = hot_local[ , j],
                                             addcovar = addcovar
                                             )
  }
  out[[j]] <- foo
}
saveRDS(out, medfile)
} else {out <- readRDS(medfile)}
```

@@

## Reading pleiotropy test results (for local-nonlocal pairs)

```{r read_pleio_results}
run_num <- 9901
pleio_file <- "_cache/Chr5_pleio.rds"

if (file.exists(pleio_file)) {mylist <- readRDS(pleio_file)} else {
  results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num)
  dir(results_dir) -> fns
  mylist <- list()
  for (i in 1:length(fns)){
    mylist[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
    as_tibble() %>%
    mutate(filename = fns[i]) %>%
    mutate(local_gene_id = stringr::str_split(fns[i], "_")[[1]][[3]], nonlocal_gene_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
    filter(lod == max(lod), trace == "profile1") %>%
    select(local_gene_id, nonlocal_gene_id, lod, filename)
  }
  saveRDS(mylist, pleio_file)
}
```

```{r read_pleio_results_local}
results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num * 10 + 1)
dir(results_dir) -> fns
pleio_file_local <- "_cache/Chr5_pleio_local.rds"
if (file.exists(pleio_file_local)) {mylist_local <- readRDS(pleio_file_local)} else {
  mylist_local <- list()
  for (i in 1:length(fns)){
    mylist_local[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
    as_tibble() %>%
    mutate(filename = fns[i]) %>%
    mutate(gene1_id = stringr::str_split(fns[i], "_")[[1]][[3]], gene2_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
    filter(lod == max(lod), trace == "profile1") %>%
    select(gene1_id, gene2_id, lod, filename)
  }
  saveRDS(mylist_local, pleio_file_local)
}

```{r read_pleio_results_nonlocal}
results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num * 10)
dir(results_dir) -> fns
pleio_file_nonlocal <- "_cache/Chr5_pleio_nonlocal.rds"
if (file.exists(pleio_file_nonlocal)) {mylist_nonlocal <- readRDS(pleio_file_nonlocal)} else {
  mylist_nonlocal <- list()
  for (i in 1:length(fns)){
    mylist_nonlocal[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
    as_tibble() %>%
    mutate(filename = fns[i]) %>%
    mutate(gene1_id = stringr::str_split(fns[i], "_")[[1]][[3]], 
           gene2_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
    filter(lod == max(lod), trace == "profile1") %>%
    select(gene1_id, gene2_id, lod, filename)
  }
  saveRDS(mylist_nonlocal, pleio_file_nonlocal)
}
```

@@

@@

## Heatmaps of pleiotropy results

### Convert pleiotropy results lists to tibbles

```{r}
local_annot3 <- local_annot2 %>%
  select(annot.id, symbol)
local_pleio <- mylist_local %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(local_annot3, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  left_join(local_annot3, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol)

nonlocal_annot3 <- nonlocal_annot %>%
  select(annot.id, symbol)
nonlocal_pleio <- mylist_nonlocal %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(nonlocal_annot3, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  left_join(nonlocal_annot3, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol)

I now bind all three tibbles that contain pleiotropy results into a single tibble.

```{r}
pleio_results_mat <- foo %>%
  dplyr::select(local_gene_id, nonlocal_gene_id, pleiotropy_lod, local_symbol, nonlocal_symbol) %>%
  rename(gene1_id = local_gene_id, 
         gene2_id = nonlocal_gene_id, 
         gene1_symbol = local_symbol,
         gene2_symbol = nonlocal_symbol
         ) %>%
  bind_rows(local_pleio, nonlocal_pleio) %>%
  hotspots::tibble_to_matrix(symmetric = TRUE)
```

```{r}
pleio_results_mat %>%
  is.na() %>%
  sum()
```

```{r}
pleio_results_mat %>%
  dist() %>%
  hclust()
```

```{r}
foobar <- pleio_results_mat %>%
  igraph::graph_from_adjacency_matrix(mode = "undirected", weighted = TRUE, diag = FALSE) %>%
  fastgreedy.community()
foobar$membership %>% table()
t1 <- tibble(gene = foobar$names, 
       community_membership = foobar$membership, 
       local = foobar$names %in% local_annot3$annot.id
       )
  table(t1$community_membership, t1$local)
```

## Clustering

```{r, hclust1}
pleio_results_mat %>%
  dist() %>%
  hclust() %>% 
  plot(cex = 0.1)
```

Check out https://genomicsclass.github.io/book/pages/clustering_and_heatmaps.html for helpful code on clustering gene expressions.

```{r rafalib_plot}
pleio_results_mat %>%
  dist() %>%
  hclust() %>% 
  rafalib::myplclust(cex = 0.1
                     )
```

```{r rafalib_plot2}
bar <- foo %>%
  filter(local_symbol == "Hnf4a")

pleio_hc <- foo %>%
  tibble_to_matrix() %>%
  t() %>%
  dist() %>%
  hclust()

rafalib::myplclust(pleio_hc,
                     lab.col = 1+ round(bar$pleiotropy_lod[match(x = bar$nonlocal_symbol, table = pleio_hc$labels)]),
                     cex = 0.3
                     )
```

```{r}
rafalib::myplclust(pleio_hc,
                   lab.col = 1 + (bar$lod_diff[match(x = bar$nonlocal_symbol,
                                                     table = pleio_hc$labels)] >
                                    mediation_threshold),
                     cex = 0.3
                     )

```{r}
library(RColorBrewer) 
library(genefilter)
library(gplots) ##Available from CRAN
cols <- palette(RColorBrewer::brewer.pal(8, "Dark2"))
#head(cbind(colnames(e),cols))
```

```{r heatmap.2}
pleio_mat <- foo %>%
  tibble_to_matrix() %>%
  t()
ld <- bar$lod_diff
idx1 <- match(x = bar$nonlocal_symbol, table = rownames(pleio_mat))
idx <- 1 + floor((ld * (ld > 0))[idx1])
cols[idx] -> indexed_colors
ic2 <- 1 + (ld[idx1] > mediation_threshold)
#indexed_colors[is.na(indexed_colors)] <- cols[8]
gplots::heatmap.2(pleio_mat, 
                  trace="none", 
                  #RowSideColors = indexed_colors, 
                  RowSideColors = c("red", "green")[ic2],
                  col=hmcol,
                  cexRow = 0.5,
                  cexCol = 0.5
                  )
```

```{r}
d3heatmap(t(pleio_mat) <= 4, cexRow = 0.2, cexCol = 0.3)
```

```{r}
gplots::heatmap.2(t(pleio_mat), 
                  trace="none", 
                  #RowSideColors = indexed_colors, 
                  ColSideColors = c("red", "green")[ic2],
                  col=hmcol,
                  cexRow = 0.5,
                  cexCol = 0.5
                  )

### 2019-07-08 analysis/paper/Chr5.Rmd
updated Rmd files for each hotspot

---
title: "Chr13 hotspot analyses"
author: "Frederick Boehm"
date: "7/08/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
hot_chr <- 13
run_num <- 9931
hot_mid <- 112.5
keller_mediator <- "Il6st"
library(hotspots)
```

>>>>

---
title: "Chr5 hotspot analyses"
author: "Frederick Boehm"
date: "7/08/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
hot_chr <- 5
run_num <- 9901
hot_mid <- 146
keller_mediator <- "Pdx1"
library(hotspots)
library(igraph)
```

### 2019-07-08 analysis/paper/Chr7.Rmd
updated Rmd files for each hotspot

---
title: "Chr5 hotspot analyses"
author: "Frederick Boehm"
date: "7/08/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
hot_chr <- 5
run_num <- 9901
hot_mid <- 146
keller_mediator <- "Pdx1"
library(hotspots)
library(igraph)
```

@@

>>>>

---
title: "Chr7 hotspot analyses"
author: "Frederick Boehm"
date: "7/08/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
hot_chr <- 7
run_num <- 9911
hot_mid <- 46
keller_mediator <- "Fam83e"
library(hotspots)
```

@@

Another possibility is to look at the local traits' pleiotropy test results only. Assign each local trait to a "pleiotropy group". Then look at each pleiotropy group, and see how similar their LOD difference values (binarized or not) are. That is, does every member of a "pleiotropy group" mediate the same set of traits?
